\bf Auxiliary Model Approach to Strongly-Correlated Systems

### 2022-03-09 notes/notes.tex
obtained upto Greens functions

\chapter{From the auxiliary model to the bulk}

@@

	\frac{1}{Nw}\sum_i \sum_{j \in \text{NN of i}} \mathcal{H}_\text{aux}(i,j) = \mathcal{H}_{H-H}

\section{Single-particle Green's function}
We now define the Greens function {\it operators}:
\begin{equation}\begin{aligned}
	\mathcal{G}_H = \frac{1}{\omega - H_{H-H}}, \mathcal{G}_\text{aux}(i,j) = \frac{1}{\omega - H_\text{aux}(i,j)}
\end{aligned}\end{equation}
Rewriting eq.~\ref{siam_to_hubb} in terms of these operators, we get
\begin{equation}\begin{aligned}
	\frac{1}{Nw}\sum_i \sum_{j \in \text{NN of i}} \left(\omega - \mathcal{G}^{-1}_\text{aux}(i,j)\right) = \omega - \mathcal{G}^{-1}_{H-H} \implies \frac{1}{Nw}\sum_i \sum_{j \in \text{NN of i}} \mathcal{G}^{-1}_\text{aux}(i,j) = \mathcal{G}^{-1}_{H-H}
\end{aligned}\end{equation}
There we used \(\frac{1}{Nw}\sum_i \sum_{j \in \text{NN of i}} \omega = \omega\).

>>>>

\chapter{From the auxiliary model to the bulk - single site approach}
	\frac{1}{N}\sum_i \mathcal{H}_\text{aux}(i) = \mathcal{H}_{H-H}

\section{Single-particle Green's function}
We now define the Greens function {\it operators}:
	\mathcal{G}_{H-H} = \frac{1}{\omega - \left(H_{H-H} - E_\text{gs}\right) }, \mathcal{G}_\text{aux}(i) = \frac{1}{\omega - \left(H_\text{aux}(i) - E_\text{gs}\right)}
Rewriting eq.~\ref{siam_to_hubb} in terms of these operators, we get
	\frac{1}{N}\sum_i \left(\omega - \mathcal{G}^{-1}_\text{aux}(i)\right) = \omega - \mathcal{G}^{-1}_{H-H} \implies \frac{1}{N}\sum_i \mathcal{G}^{-1}_\text{aux}(i) = \mathcal{G}^{-1}_{H-H}~,
where we used \(\frac{1}{N}\sum_i \omega = \omega\). We will now take matrix elements of the full Greens function and obtain these matrix elements in terms of those of the auxiliary model. Let \(\left\{ \ket{\Phi}_n \right\} \) be the set of eigenstates of \(H_\text{aux}\), and \(\ket{\Phi}_0\) be the groundstate. We assume that the ground state of \(H_{H-H}\) is captured well by the auxiliary model, such that the real-space diagonal matrix element for particle propagation is obtained by sandwiching the Greens function between the states \(c^\dagger_{i\sigma}\ket{\Phi}_0\):
	\left(\mathcal{G}^{-1}_{H-H}\right)^p_{ii} \equiv \braket{\Phi_0 | c_{i\sigma} \mathcal{G}^{-1}_{H-H} c^\dagger_{i\sigma} | \Phi_0} = \frac{1}{Nw}\sum_j \sum_{k \in \text{NN of j}}\braket{\Phi_0 | c_{i\sigma} \mathcal{G}^{-1}_\text{aux}(j,k) c^\dagger_{i\sigma} | \Phi_0}
The superscript \(p\) indicates that this is the particle propagation matrix element.
Under the single site approximation, we ignore the hopping between multiple impurity that reside in separate instances of the auxiliary model. Therefore, the only operator \(\mathcal{G}^{-1}_\text{aux}(j)\) that affects the matrix element is \(j=i, k \in \text{NN of i}\). Moreover, because of translation invariance, all \(k\) are equivalent.
	\left(\mathcal{G}^{-1}_{H-H}\right)^p_{ii} = \frac{1}{Nw} w\times \braket{\Phi_0 | c_{i\sigma} \mathcal{G}^{-1}_\text{aux}(i) c^\dagger_{i\sigma} | \Phi_0} = \frac{1}{N}\braket{\Phi_0 | c_{d\sigma} \mathcal{G}^{-1}_\text{aux}(d) c^\dagger_{d\sigma} | \Phi_0}
where we have identitied site \(i\) in the auxiliary model as the impurity.
By inserting \(1 = \sum_n \ket{\Phi_n}\bra{\Phi_n}\) on both sides of \(\mathcal{G}^{-1}_\text{aux}(d)\), we get
	\left(\mathcal{G}^{-1}_{H-H}(\omega)\right)^p_{ii} = \frac{1}{N}\sum_{m,n} \braket{\Phi_0 | c_{d\sigma} \ket{\Phi_m}\bra{\Phi_m}\mathcal{G}^{-1}_\text{aux}(d) \ket{\Phi_n}\bra{\Phi_n} c^\dagger_{d\sigma} | \Phi_0} = \left(\mathcal{G}^{-1}_{H-H}\right)_{ii} = \frac{1}{N}\sum_{n} |d^p_n|^2 \left(\mathcal{G}^{-1}_\text{aux}(d, \omega)\right)_{nn} 
There we used the fact that since \(\ket{\Phi_m}\) are eigenstates of \(H_\text{aux}\) and hence of \(\mathcal{G}^{-1}_\text{aux}\), we can set \(m=n\). We also defined \(d^p_n = \bra{\Phi_n}c^\dagger_{d\sigma}\ket{\Phi_0}\). The hole propagation matrix element can be obtained similarly:
	\left(\mathcal{G}^{-1}_{H-H}(-\omega)\right)^h_{ii} = \frac{1}{N}\sum_{n} |d^h_n|^2 \left(\mathcal{G}^{-1}_\text{aux}(d, -\omega)\right)_{nn} 
where \(d^h_n = \bra{\Phi_n}c_{d\sigma}\ket{\Phi_0}\). The off-diagonal matrix elements can also be obtained similarly. The difference from the diagonal case is that only one out of the \(w\) nearest-neighbour pairs contributes to the matrix element:
	\left(\mathcal{G}^{-1}_{H-H}\right)^p_{i,i+1} \equiv \braket{\Phi_0 | c_{i,\sigma} \mathcal{G}^{-1}_{H-H} c^\dagger_{i+1,\sigma} | \Phi_0} = \frac{1}{Nw}\braket{\Phi_0 | c_{i\sigma} \mathcal{G}^{-1}_\text{aux}(i,i+1) c^\dagger_{i+1,\sigma} | \Phi_0} = \frac{1}{Nw}\braket{\Phi_0 | c_{d\sigma} \mathcal{G}^{-1}_\text{aux}(d) c^\dagger_{z,\sigma} | \Phi_0}
Here, we identity \(i+1\) as the bath zero mode \(z\) and \(i\) as the impurity \(d\). We again define matrix elements \(z_n^p = \braket{\Phi_n | c^\dagger_{z\sigma} | \Phi_0}, z_n^h = \braket{\Phi_n | c_{z\sigma} | \Phi_0}\). The matrix element becomes
	\left(\mathcal{G}^{-1}_{H-H}\left(\omega\right) \right)^p_{i,i+1} = \frac{1}{Nw} \sum_n \left(d^p_n\right)^* z^p_n \left(\mathcal{G}^{-1}_\text{aux}(d, \omega) \right)_{nn}
The hole counterpart is
	\left(\mathcal{G}^{-1}_{H-H}\left(-\omega\right) \right)^h_{i,i+1} = \frac{1}{Nw} \sum_n \left(d^h_n\right)^* z^h_n \left(\mathcal{G}^{-1}_\text{aux}(d, -\omega) \right)_{nn}
In summary, the four Greens functions are
	\left(\mathcal{G}^{-1}_{H-H}(\omega)\right)^p_{ii} = \frac{1}{N}\sum_{n} |d^p_n|^2 \left(\mathcal{G}^{-1}_\text{aux}(d, \omega)\right)_{nn}, ~ ~ ~ \left(\mathcal{G}^{-1}_{H-H}(-\omega)\right)^h_{ii} = \frac{1}{N}\sum_{n} |d^h_n|^2 \left(\mathcal{G}^{-1}_\text{aux}(d, -\omega)\right)_{nn}\\
	\left(\mathcal{G}^{-1}_{H-H}\left(\omega\right) \right)^p_{i,i+1} = \frac{1}{Nw} \sum_n \left(d^p_n\right)^* z^p_n \left(\mathcal{G}^{-1}_\text{aux}(d, \omega) \right)_{nn}, ~ ~ ~ \left(\mathcal{G}^{-1}_{H-H}\left(-\omega\right) \right)^h_{i,i+1} = \frac{1}{Nw} \sum_n \left(d^h_n\right)^* z^h_n \left(\mathcal{G}^{-1}_\text{aux}(d, -\omega) \right)_{nn}

In order to obtain the matrix elements of \(\mathcal{G}_{H-H}\) as compared to its inverse, we note that the matrix elements of the inverse are related to the matrix elements of another inverse operator, through some coefficients. We can therefore use the identity
\begin{equation}\begin{aligned}
\left(\hat O\right)_{ij} = \bra{i} \hat O \ket{j} = \sum_{m,n} \braket{i|m} \left(\hat O\right)_{mn} \braket{n|j} \implies \left(\hat O^{-1}\right)_{ij} = \sum_{m,n} \braket{i|m} \left(\hat O^{-1}\right)_{mn} \braket{n|j}
\end{aligned}\end{equation}
We can identity \(\left(\hat O^{-1}\right)_{mn}\) as the matrix elements of \(\mathcal{G}_\text{aux}^{-1}\) in our expressions. Using the relation, we can therefore write
\begin{gather}
	\left(\mathcal{G}_{H-H}(\omega)\right)^p_{ii} = \frac{1}{N}\sum_{n} |d^p_n|^2 \left(\mathcal{G}_\text{aux}(d, \omega)\right)_{nn}, ~ ~ ~ \left(\mathcal{G}_{H-H}(-\omega)\right)^h_{ii} = \frac{1}{N}\sum_{n} |d^h_n|^2 \left(\mathcal{G}_\text{aux}(d, -\omega)\right)_{nn}\\
	\left(\mathcal{G}_{H-H}\left(\omega\right) \right)^p_{i,i+1} = \frac{1}{Nw} \sum_n \left(d^p_n\right)^* z^p_n \left(\mathcal{G}_\text{aux}(d, \omega) \right)_{nn}, ~ ~ ~ \left(\mathcal{G}_{H-H}\left(-\omega\right) \right)^h_{i,i+1} = \frac{1}{Nw} \sum_n \left(d^h_n\right)^* z^h_n \left(\mathcal{G}_\text{aux}(d, -\omega) \right)_{nn}
\end{gather}

The single-particle Greens functions can now be written in terms of these matrix elements. Using eq.~\ref{G_mat_el}, we write
\begin{equation}\begin{aligned}
	\left(G_{H-H}(\omega)\right)_\text{loc} &= \left(G_{H-H}(\omega)\right)_{ii}^p - \left(G_{H-H}(-\omega)\right)_{ii}^h = \frac{1}{N}\sum_n \left[|d^p_n|^2 \left(\mathcal{G}_\text{aux}(d, \omega)\right)_{nn} - |d^h_n|^2 \left(\mathcal{G}_\text{aux}(d, -\omega)\right)_{nn}\right] \\
	\left(G_{H-H}(\omega)\right)_\text{nn} &= \left(G_{H-H}(\omega)\right)_{i,i+1}^p - \left(G_{H-H}(-\omega)\right)_{i+1,i}^h = \frac{1}{Nw}\sum_n \left[\left(d^p_n\right)^* z^p_n \left(\mathcal{G}_\text{aux}(d, \omega) \right)_{nn} - \left(z^h_n\right)^* d^h_n \left(\mathcal{G}_\text{aux}(d, -\omega)\right)_{nn}\right] 
\end{aligned}\end{equation}

\chapter{From the auxiliary model to the bulk - two site approach}

@@
